---
title: "End-to-End Machine Learning with H2O DSC5.0 Tutorial, Part 2 (bit.ly/dsc50_h2o_tutorial)"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    number_sections: yes
    fig_height: 10
    fig_width: 14
    highlight: kate
    toc_depth: 3
    css: style.css
    
---


```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)
library(DT)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Classification Part One: H2O AutoML

```{r, message=FALSE}
# Let's go
library(h2o) # for H2O Machine Learning
library(mlbench) # for Datasets
```

```{r}
# Enter your lucky seed here ...
n_seed <- 12345
```

## Data - Pima Indians Diabetes from `mlbench`

```{r}
data("PimaIndiansDiabetes")
datatable(head(PimaIndiansDiabetes), 
          rownames = FALSE, options = list(pageLength = 6, scrollX = TRUE))
```


## Data Prep

```{r}
# Convert pos and neg to 1 and 0
d_new <- PimaIndiansDiabetes[, -ncol(PimaIndiansDiabetes)]
d_new$diabetes <- 0
d_new[which(PimaIndiansDiabetes$diabetes == "pos"), ]$diabetes <- 1
PimaIndiansDiabetes <- d_new
rm(d_new)
```

```{r}
target <- "diabetes" 
features <- setdiff(colnames(PimaIndiansDiabetes), target)
print(features)
```



## Start a local H2O Cluster (JVM)

```{r}
h2o.init()
```

```{r}
h2o.no_progress() # disable progress bar for RMarkdown
h2o.removeAll()   # Optional: remove anything from previous session 
```



## Convert R dataframe into H2O dataframe

```{r}
# H2O dataframe
h_diabetes <- as.h2o(PimaIndiansDiabetes)

# Make sure the target is a factor (for classification)
h_diabetes$diabetes <- as.factor(h_diabetes$diabetes)
```



## Split Data into Train/Test

```{r}
h_split <- h2o.splitFrame(h_diabetes, ratios = 0.8, seed = n_seed)
h_train <- h_split[[1]] # 80% for modelling
h_test <- h_split[[2]] # 20% for evaluation
```

```{r}
dim(h_train)
dim(h_test)
```


## H2O AutoML

```{r}
# Run AutoML (try n different models)
# Check out all options using ?h2o.automl
automl = h2o.automl(x = features,
                    y = target,
                    training_frame = h_train,
                    nfolds = 5,                        # 5-fold Cross-Validation
                    max_models = 20,                   # Max number of models
                    stopping_metric = "logloss",       # Metric to optimize
                    project_name = "automl_diabetes",  # Specify a name so you can add more models later
                    sort_metric = "logloss",
                    seed = n_seed)
```

### Leaderboard

```{r}
datatable(as.data.frame(automl@leaderboard), 
          rownames = FALSE, options = list(pageLength = 10, scrollX = TRUE, round)) %>%
  formatRound(columns = -1, digits = 4)
```

# Classification Part Two: XAI

## Package `DALEX`

```{r}
# Descriptive mAchine Learning EXplanations (DALEX)
library(DALEX)
```


### The `explain()` Function

```{r}
# Custom Predict Function
custom_predict <- function(model, newdata) {
  newdata_h2o <- as.h2o(newdata)
  res <- as.data.frame(h2o.predict(model, newdata_h2o))
  return(round(res$p1)) # round the probabil
  }
```

### Explainer for H2O Models

```{r}
explainer_automl <- DALEX::explain(model = automl@leader, 
                                data = as.data.frame(h_test)[, features],
                                y = as.numeric(as.character(as.data.frame(h_test)[, target])),
                                predict_function = custom_predict,
                                label = "H2O AutoML")
```

### Variable importance

```{r}
library(ingredients)
vi_automl <- feature_importance(explainer_automl, type="difference")
plot(vi_automl)
```


### Partial Dependence Plots

Let's look at feature `age` 

```{r}
pdp_automl_rm <- partial_dependency(explainer_automl, variables = "age")
plot(pdp_automl_rm)
```


## Prediction Understanding

```{r}
library(iBreakDown)
```

```{r}
# Prediction: Diabetes = Negative (0)
pb_automl <- break_down(explainer_automl, new_observation = as.data.frame(h_test)[1, ])
plot(pb_automl)
```


```{r}
# Prediction: Diabetes = Positive (1)
pb_automl <- break_down(explainer_automl, new_observation = as.data.frame(h_test)[6, ])
plot(pb_automl)
```


# Bring Your Own Data + Q&A

Get your hands dirty!

